### 2017-07-31 README.Rmd
add scripts for using toy data
add bib for readme refs

---
output: github_document
---

@@

`methyvim` is an R package that provides facilities for differential methylation
analysis based on _variable importance measures_ (VIMs), a class of statistical
target parameters that arise in causal inference. The statistical techniques
implemented rely on the use of targeted minimum loss-based estimation (TMLE) to
assess two particular VIMs: (1) the "local" average treatment effect for
discrete exposures, and (2) a nonparametric variable importance measure for
continuous exposures. Within this framework, these methods allow differential
methylation effects to be estimated in an assumption-free manner, on a
pre-screened set of genomic sites measured by DNA methylation assays. As the
statistical algorithm uses a multi-stage approach, multiple testing corrections
are made using a modified marginal Benjamini & Hochberg step-up False Discovery
Rate controlling procedure for multi-stage analyses (FDR-MSA). In order to allow
the user significant flexibility with respect to the scientific questions posed,
the procedure estimates one of the appropriate VIMs for each of a reduced set of
genomic sites, using screening procedures to identify a reduced set from the
full data when a reduced set is not pre-specified. While making allowance for
the user to specify the reduced set of genomic sites to be tested (e.g., those
related to a particular gene of interest), this package comes equipped with
screening procedures to facilitate the initial reduction of genomic sites, the
two most noteworthy of these being a procedure based on the well-known R package
[`limma`](https://bioconductor.org/packages/release/bioc/html/limma.html) and an
extension of data-adaptive statistical target parameters for multiple testing
(implemented in the R package
[`DA.Test`](https://github.com/wilsoncai1992/data.adapt.multi.test)).

@@

To contribute, install the _development version_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-develop-installation, eval = FALSE}
devtools::install_github("nhejazi/methyvim", ref = "develop")
```

@@

After using the `methyvim` R package, please cite it:

        @article{hejazi2017methyvim,
          doi = {},
          url = {},
          year  = {2017},
          month = {},
          publisher={Faculty of 1000 Ltd},
          volume = {},
          author = {Hejazi, Nima S. and Hubbard, Alan E. and {van der Laan},
                    Mark J.},
          title = {methyvim: Nonparametric Assessment of Variable Importance for
                   Differential Methylation Analysis},
          journal = {F1000Research}
        }

@@

* [Mark J. van der Laan & Sherri Rose. _Targeted Learning: Causal Inference for
    Observational and Experimental Data_,
    2011.](http://www.targetedlearningbook.com)

* [Antoine Chambaz, Pierre Neuvial, & Mark J. van der Laan. "Estimation of a
    non-parametric variable importance measure of a continuous exposure",
    _Electronic Journal of Statistics_,
    2012.](http://www.math-info.univ-paris5.fr/~chambaz/Papiers/chambazNeuvialvanderLaan_EJS2012.pdf)

* [Catherine Tuglus & Mark J. van der Laan. "FDR controlling procedures for
    multi-stage analyses", _UC Berkeley Division of Biostatistics Working
    Paper Series_, 2008.](http://biostats.bepress.com/ucbbiostat/paper239/)

* [Weixin Cai, Nima S. Hejazi, & Alan E. Hubbard. "Data-adaptive statistics for
    multiple testing in high-dimensional settings", _In preparation_,
    2017.](https://www.overleaf.com/5660573pjjrxh#/25678897/)

* [Gordon K. Smyth. "Linear models and empirical Bayes methods for assessing
    differential expression in microarray experiments." _Statistical
    Applications in Genetics and Molecular Biology_, 3(1),
    2004.](http://www.statsci.org/smyth/pubs/ebayes.pdf)

>>>>

---
output: github_document
bibliography: refs.bib
---

@@

## Description

`methyvim` is an R package that provides facilities for differential methylation
analysis based on _variable importance measures_ (VIMs), a class of statistical
target parameters that arise in causal inference.

The statistical methodology implemented computes targeted minimum loss-based
estimates of well-studied variable importance measures:

1. The average treatment effect (ATE) for discrete exposures/treatments,
2. A nonparametric variable importance measure (NPVI) for continuous exposures
[@chambaz2012estimation].

These methods allow differential methylation effects to be quantified in a
manner that is largely free of assumptions, especially of the variety exploited
in parametric models. __The statistical algorithm consists in several major
steps:__

1. Pre-screening of genomic sites is used to isolate a subset of sites for
which there is evidence of differential methylation. Targeted minimum
loss-based estimates of VIMs are computed for this subset of sites only (for
the sake of computational feasibility). Several screening approaches are
available, adapting core routines from the following R packages:
[`limma`](http://bioconductor.org/packages/release/bioc/html/limma.html),
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi),
[`adaptest`](https://github.com/nhejazi/adaptest).
2. Nonparametric VIMs are estimated for the specified parameter, currently using
routines from the R packages
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi) to estimate the
NPVI parameter, or [`tmle`](https://CRAN.R-project.org/package=tmle) to
estimate the ATE.
3. Since pre-screening is performed prior to estimating VIMs, we make use of a
multiple testing correction uniquely suited to such settings, specifically we
use the modified marginal Benjamini & Hochberg step-up False Discovery Rate
controlling procedure for multi-stage analyses (FDR-MSA) [@tuglus2009modified].

See @vdl2011targeted for a general description of the framework of targeted
minimum loss-based estimation.

Note about shrinkage of influence curves, adapting @smyth2004linear.

More exposition here....

@@

<!--
To contribute, install the _development version_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-develop-installation, eval = FALSE}
devtools::install_github("nhejazi/methyvim", ref = "develop")
```
-->

@@

After using the `methyvim` R package, please cite both of the following:

        @article{hejazi2017methyvim,
          doi = {},
          url = {},
          year  = {2017},
          month = {},
          publisher={Faculty of 1000 Ltd},
          volume = {},
          author = {Hejazi, Nima S. and Hubbard, Alan E. and {van der Laan},
            Mark J.},
          title = {methyvim: Differential methylation analysis with targeted
            nonparametric variable importance measures},
          journal = {F1000Research}
        }

        @misc{hejazi2017methyvim-r,
          doi = {},
          url = {},
          howpublished = {\url{https://github.com/nhejazi/methyvim}},
          year  = {2017},
          author = {Hejazi, Nima S. and Hubbard, Alan E. and {van der Laan},
            Mark J.},
          title = {methyvim: Differential methylation analysis with targeted
            minimum loss-based estimates of variable importance measures}
        }

@@
